---
title: "Homework Week 2"
author: "Eric Pitruzzella epitruzz@gmail.com"
date: "September 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

##Solution
Set up Connetion to MySQL 

```{r}
p=getPass::getPass()

library(RMySQL)

mydb = dbConnect(MySQL(), user='epitruzz', password=p, dbname='movie_ratedb', host='localhost')

```
Created 3 tables in MYSQL earlier, here they are
```{r}
dbListTables(mydb)

```
Read all data from MySQL of all crtics and their rating
```{r}
myfilms <- dbSendQuery(mydb, "SELECT Replace(a.name,'\r','') as 'Critic', 
Replace(f.name,'\r','') as 'Movie', 
r.rating 
from filmRating r, films f, raters a
where r.id=f.id and r.nid=a.nid order by a.name;")
dbFetch(myfilms)
```
Show each movie's average rating
```{r}
myfilms <- dbSendQuery(mydb, "SELECT Replace(f.name,'\r','')as 'Movie',avg(r.rating) as 'Avg Rating' from filmRating r, films f, raters a
where r.id=f.id and r.nid=a.nid group by f.name order by avg(r.rating) desc ;")
dbFetch(myfilms)
```
Close Connections
```{r}
 dbDisconnect(mydb)
```





