---
title: "Data 607 hw3"
author: "Eric Pitruzzella"
date: "September 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

3. Copy the introductory example. The vector name stores the extracted names.
R> name
[1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
[4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"

```{r}
#rm(list=ls())  #clear everything
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
```


(a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

load into name vector

```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
```

Remove titles and initials. Trim whitespaces
```{r}
name2 <- str_trim(sub("[[:alpha:]]{1,3}\\.", '', name))
name2
```

Replace instances of "Last name comma first name"  with "first name lastname" by Backreferencing.

```{r}
name3 <- sub("(\\w+),\\s+(\\w+)","\\2 \\1", name2)
name3
```

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).
```{r}
nameTitle <- str_detect(name,"Rev|Dr[.]")
nameTitle
```

(c) Construct a logical vector indicating whether a character has a second name.
```{r}
namecount <- str_count(str_trim(sub("Rev|Dr[.]", "", name)),"\\w+")
secndname <- str_detect(namecount, "3") # anything with more than 3 words has a second name
secndname
```

4.Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.

(a) [0-9]+\\$

a string of numbers followed by a $
```{r}
pattern="[0-9]+\\$"
ex <- "1234$ ab$a 1$ 00012324"
str_extract_all(ex, pattern)

```
(b) \\b[a-z]{1,4}\\b

get any word (as separated by the word barriers \\b)all in lowercase letters. Words matched are between 1 and 4 characters long
```{r}
ex1 = "How do I know the things"
str_extract_all(ex1,"\\b[a-z]{1,4}\\b")
```

(c) .*?\\.txt$
 any strings that end in .txt 
 
```{r}
ex2 = "filename.dat something.txt"
str_extract_all(ex2, ".*?\\.txt$")
```
(d) \\d{2}/\\d{2}/\\d{4}

Any values that have 2 numbers slash 2 numbers slash 4 numbers. For example a date. 
```{r}
date1 <- c("12/25/1978 2017/16/12 99/99/1978")
str_extract_all(date1, "\\d{2}/\\d{2}/\\d{4}")
```
(e) <(.+?)>.+?</\\1>
Any string inside < > followed by any string and followed by the same values inside the previous < > except ending with </ >.
```{r}
ex3 = "<html>    <head>        <title>CUNY - Self Service</title>"
str_extract_all(ex3, "<(.+?)>.+?</\\1>")
```

9. The following code hides a secret message. Crack it with R and regular expressions.
Hint: Some of the characters are more revealing than others! The code snippet is also
available in the materials at www.r-datacollection.com.
clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr

Extract all the uppercase letters from the message and concatenate into a single vector 
```{r}
x="clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
str_c(unlist(str_extract_all(x, "[[:upper:].]")),collapse = "")
```

