---
title: "Week 5 Assignment"
author: "Eric Pitruzzella"
date: "September 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load.packages <- c("stringr","prettydoc", "tidyr", "dplyr", "knitr")
```

Load Packages
```{r}

library(tidyr)
library(dplyr)
library(stringr)
```

(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

Add headers for Airline and status and row headers for "delayed".

```{r}
csv_file <- rbind(c("Airline", "Status", "Los Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle"),
             c("ALASKA", "On Time", 497, 221, 212, 503, 1841),
             c("ALASKA", "Delayed", 62, 12, 20, 102, 305),
             c("AM WEST", "On Time", 694, 4840, 383, 320, 201),
             c("AM WEST", "Delayed", 117, 415, 65, 129, 61))


csv_file
write.table(csv_file, file = "Airline.csv", sep = ",", col.names=F, row.names=F)
```

(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. 

Read the data from the file

```{r}
csvdata <- read.csv(paste0("Airline.csv"), stringsAsFactors = F);
csvdata[,2] <- sapply(csvdata[,2], str_replace, " ", "_")
csvdata
```


Use the gather function was used to move data from wide to tall format.

```{r}
csv_file2 <- gather(csvdata, "City", "Count", 3:7)
csv_file2
```

Use the spread function to move Status into its own variable

```{r}
csv_file2 <- spread(csv_file2, Status, Count)
csv_file2
```

(3) Perform analysis to compare the arrival delays for the two airlines.

examine delays for these two airlines in general and by city

By Airline:
```{r}
csv_file2 %>% group_by(Airline) %>%
summarise(
TotDelay=sum(Delayed),TotOn_Time=sum(On_Time), 
PctDelayed=(TotDelay/(TotDelay+TotOn_Time)*100),
PctOnTime=(TotOn_Time/(TotDelay+TotOn_Time)*100)
)
```
Alaska has a greater percentage of delays than AM West. 


By City:
```{r}
csv_file2 %>% group_by(City) %>%
summarise(
TotDelay=sum(Delayed),TotOn_Time=sum(On_Time), 
PctDelayed=(TotDelay/(TotDelay+TotOn_Time)*100),
PctOnTime=(TotOn_Time/(TotDelay+TotOn_Time)*100)
)
```
San Francisco has the greatest  percentage of delays of the cities listed. Phoenix has the best on time. 


By City and Airline:
```{r}
csv_file2 %>% group_by(City,Airline) %>%
summarise(
TotDelay=sum(Delayed),TotOn_Time=sum(On_Time), 
PctDelayed=(TotDelay/(TotDelay+TotOn_Time)*100),
PctOnTime=(TotOn_Time/(TotDelay+TotOn_Time)*100)
)
```
In particular, AM West at San Francisco has a much higher percentage of delays than all airlines and cities. Taking Alaska airlines to Phoenix would have the best chance of getting there on time. 



---


